Ultra low thermal conductivity of quasi 1D bulk material \bii

### 2017-06-30 latex/paper.tex
paper

% @Author: YangZhou
% @Date:   2017-06-20 20:27:38
% @Last Modified by:   YangZhou
% @Last Modified time: 2017-06-21 02:50:42
\documentclass[%
 reprint,
 amsmath,amssymb,
 aps,
prb,
]{revtex4-1}
\preprint{APS/123-QED}
\usepackage{graphicx}% Include figure files
\usepackage{bm}% bold math
\newcommand{\angstrom}{\mbox{\normalfont\AA}}
\newcommand{\bii}{$\beta-Bi_4 I_4$\;}
\begin{document}
\title{Ultra low thermal conductivity of quasi 1D bulk material \bii}
\author{Yang Zhou}
 \altaffiliation[Also at ]{Physics Department, Fudan University.}%Lines break automatically or can be forced with \\
\author{Xin-Gao Gong}%
 \email{xggong@fudan.edu.cn}
\affiliation{%
 Authors' institution and/or address\\
 This line break forced with \textbackslash\textbackslash
}%
\begin{abstract}
The thermal transportation properties of recently synthesized quasi-1D bulk \bii  is explored from first principle. The thermal conductivity of this system is found rather low because of the phonon localization caused by the low-dimension nature and large anharmonic scattering of phonons. Detailed analysis of lattice dynamic calculation is applied to explain the origin of the anharmonic effects. The Seebeck coefficients and electric conductance are calculated from first principle and then ZT value is given. We found \bii  is actually potential for the application of thermoelectric. This research could be a guidance for further exploring of advanced low thermal conductivity materials.
\begin{description}
\item[Keywords]
Thermal conductivity, quasi-1D,\bii, thermoelectric, phonons
\end{description}
\end{abstract}

@@

Thermoelectric (TE) materials can directly convert heat into electricity or vice versa and, consequently, have attracted wide attention for applications in energy harvesting, and solid state cooling that does not require refrigerants \cite{Majumdar2004,Zebarjadi2012}. The figure of merit for thermoelectric performance is $ZT= S^2 \sigma T/(\kappa_L+\kappa_el)$ , where S and $\sigma$ are the Seebeck coefficient and electrical conductivity, and $\kappa_L$ and $\kappa_el$ are the lattice and electronic thermal conductivities, respectively. Controlling of the power factor and the reduction of thermal conductivity due to phonon scattering are available ways to overcome the limits of the current ZT value. However, it is very difficult to significantly improve the thermoelectric performance of conventional materials because the $\sigma$, S and $\kappa_L$ are coupled unfavorably. For instance, a high electrical conductivity usually leads to a low Seebeck coefficient and a high electrical thermal conductivity, both of which are undesirable for thermoelectric applications.

@@

@@

\begin{figure}[b]
\includegraphics[width=0.4\textwidth]{images/structure}
\caption{\label{fig:structure} (color online) Crystal structure of \bii. a) the perspective view of the bulk structure. Bi atoms are in blue and I atoms in purple. The crystal axes in the structure are chosen to match the stripe direction, in order to facilitate the calculation of thermal conductivity along the stripe. b) the side view of an individual chain-like stripe.}
\end{figure}

@@

\begin{figure}[b]
\includegraphics[width=0.4\textwidth]{images/phonon}
\caption{\label{fig:phonon} (color online) Phonon dispersion along the high symmetric k-point path in the first Brillouin zone, along with phonon density of states.}
\end{figure}

To further study its thermal stability at finite-temperature, we performed ab initio molecular dynamics (MD) simulations at typical temperatures and the results of Mean Square Displacement(MSD)\cite{Michalet2010} are shown in FIG.3. The MSD is defined as

@@

@@

However, the structure starts melting when heated to 900 K. Thus we only concentrated on the trusted temperature range (300¨C800 K), and choose 300 K as a typical temperature to perform the thermal conductivity calculations.

To obtain reliable results of thermal conductivity, the convergence dependence to q points density has to be tested. FIG.4a shows the convergence of thermal conductivity on $64 \times N \times N$ q points with N increase from 2 to 8. The TC shows its convergence at N=4, above witch the result changes a little but not the computational effort, so we choose the number of transverse q points to be both 4. Besides, the convergence on $N_x \times 4 \times 4$ q points with N increase from 16 to 1024 is also tested as in FIG.4b. The result of $N_x=200$ could be a good estimate of the converged result and we choose it to calculate TC only for the trends of TC on other properties like temperature et. al to save computer resources. FIG.4c shows the convergence on phonon cut off free path. The exact convergence could not be obtained even at $N_x=1024$ due to the abnormally large lifetime of long wave phonons show in FIG.4d, which always exist in low dimensional materials. However, a lot of research shows the relation of TC and phonon cut off free path obeys the rules of \cite{Schelling2002}

@@

and FIG.4e verifies that it also make sense in bulk \bii and the calculated results are over estimated. As the conclusion, the calculated results of thermal conductivity are not converged even in extremely dense q points, however they can be used as upper limited estimates of the exact results.

The thermal conductivity of \bii is extremely low and anisotropic as shown in FIG.5. The TC values decreases with increasing temperature and obeys the rule of $¦Ê~1/T$, and the largest value at 300K is much less than $1 W/mK$. The TC along the stripes are almost 3 times that of the transverse value which may result from the nature the quasi-1D structure.

@@

where N is the total number of atoms and $\epsilon_{k\sigma} (s)$ is the complex amplitude of atom s for eigen-mode k¦Ò. The participation ratio presents the fraction of atoms participating in a given mode and effectively indicates the localized modes with $O(1/N)$ and delocalized modes with $O (1)$. It can provide a more detailed information about the localization effect to each mode. The eigenvectors and frequencies are obtained using Phonopy\cite{Togo2008} with $3\times2\times2$ supercell and $15\times15\times15$ mesh sampling. FIG.8 shows the participation ratio of \bii. A lot of phonons have participation ratio below 0.5 which means they are strongly localized and lost the capability to carry energy and they have little contribution to thermal conductivity.

Large anharmonic would be another key factor that account for the low thermal conductivity. FIG.6 shows the group velocity of the phonons. Although the frequencies are pretty low (phonons are pretty soft), the group velocities are not so small witch can be understood by the steep phonon dispersion, for example, along the path $\Gamma M$. The maximum group velocity is comparable to that of graphene witch possesses the largest thermal conductivity so far. So group velocity is not responsible for the low TC. The large Gruneisen parameters\cite{Hasegawa1980} showed in FIG.9b means pretty large anharmonic effects in \bii which suggest strongly three-phonon scattering effects\cite{Hungary1985Phonon}. FIG.7a shows the three-phonon scattering relaxation time of all the phonons. The mean free path are also show in FIG.7b.Except for the long wave ones, all the life time are pretty small and this is exactly the origin of the low TC. The scattering strength show in FIG.9a also shows large scattering of phonons.

Low TC means the potential of applications in thermoelectric field and \bii is predicted to be a good thermoelectric material. FIG.10a shows Seebeck coefficient S at different temperatures. The typical values are in the range of several mV/K and are comparable with that of SnSe. Seebeck coefficient are sensitive to temperature while FIG.10b shows that electric conduction $\sigma$ are not. The power factor $S^2 \sigma$ are shown in FIG.10c. Although Seebeck coefficient is smaller at 700K, power factor is higher because of the rise and fall of $\sigma$ around fermi level. However $\kappa_el$ is also higher at 700K which eliminate the difference. The electric thermal conductivity is comparable at all the temperature and is even higher, which dominant at high temperature as shown in FIG.11a. At this point all the factors of figure ZT of merit are collected and the final results are shown in FIG.11b. After some n-doping or p-doping, ZT reaches its maximum and the best values could be obtained at high temperature. At 700K this ZT value is around 0.8 and this is not bad for the application of thermoelectric energy translation.

>>>>

% @Author: YangZhou
% @Date:   2017-06-20 20:27:38
% @Last Modified by:   YangZhou
% @Last Modified time: 2017-06-30 17:13:36
\documentclass[%
 reprint,
 amsmath,amssymb,
 aps,
 prb,
]{revtex4-1}
\preprint{APS/123-QED}
\usepackage{graphicx}% Include figure files
\usepackage{bm}% bold math
\newcommand{\angstrom}{\mbox{\normalfont\AA}}
\newcommand{\bii}{$\beta-Bi_4 I_4$\;}
\begin{document}
\title{Ultra low thermal conductivity of quasi 1D bulk material \bii}
\author{Yang Zhou${}^{1,2}$}
\author{Zhi-Xin Guo${}^{3}$}
\author{Shi-you Chen}
\author{Hongjun Xiang}
\author{Xin-Gao Gong${}^{1,2}$}
\email{xggong@fudan.edu.cn}
\affiliation{%
  ${}^1$Key Laboratory for Computational Physical Science (Ministry of Education), State Key Laboratory of Surface Physics and Department of Physics, Fudan University, Shanghai 200433, China\\
  ${}^2$Collaborative Innovation Center of Advanced Microstructures, Nanjing 210093, Jiangsu, China\\
  ${}^3$Department of Physics, Xiangtan University, Xiangtan 411105, China
}%
\begin{abstract}
  The thermal transportation properties of recently synthesized quasi-1D bulk \bii  is explored from first principle. The thermal conductivity of this system is found rather low because of the phonon localization caused by the low-dimension nature and large anharmonic scattering of phonons. Detailed analysis of lattice dynamic calculation is applied to explain the origin of the anharmonic effects. The Seebeck coefficients and electric conductance are calculated from first principle and then ZT value is given. We found \bii  is actually potential for the application of thermoelectric. This research could be a guidance for further exploring of advanced low thermal conductivity materials.
  \begin{description}
    \item[Keywords]
          Thermal conductivity, quasi-1D,\bii, thermoelectric, phonons
  \end{description}
\end{abstract}

@@

Thermoelectric (TE) materials can directly convert heat into electricity or vice versa and, consequently, have attracted wide attention for applications in energy harvesting, and solid state cooling that does not require refrigerants \cite{Majumdar2004,Zebarjadi2012}. The figure of merit for thermoelectric performance is $ZT= S^2 \sigma T/(\kappa_L+\kappa_{el})$ , where S and $\sigma$ are the Seebeck coefficient and electrical conductivity, and $\kappa_L$ and $\kappa_{el}$ are the lattice and electronic thermal conductivities, respectively. Controlling of the power factor and the reduction of thermal conductivity due to phonon scattering are available ways to overcome the limits of the current ZT value. However, it is very difficult to significantly improve the thermoelectric performance of conventional materials because the $\sigma$, S and $\kappa_L$ are coupled unfavorably. For instance, a high electrical conductivity usually leads to a low Seebeck coefficient and a high electrical thermal conductivity, both of which are undesirable for thermoelectric applications.

@@

\begin{figure}[b]
  \includegraphics[width=0.48\textwidth]{images/structure}
  \caption{\label{fig:structure} (color online) Crystal structure of \bii. a) the perspective view of the bulk structure. Bi atoms are in blue and I atoms in purple. The crystal axes in the structure are chosen to match the stripe direction, in order to facilitate the calculation of thermal conductivity along the stripe. b) the side view of an individual chain-like stripe.}
\end{figure}

@@

@@

\begin{figure}
  \includegraphics[width=0.48\textwidth]{images/phonon}
  \caption{\label{fig:phonon} (color online) Phonon dispersion along the high symmetric k-point path in the first Brillouin zone, along with phonon density of states.}
\end{figure}

To further study its thermal stability at finite-temperature, we performed ab initio molecular dynamics (MD) simulations at typical temperatures and the results of Mean Square Displacement(MSD)\cite{Michalet2010} are shown in FIG.\ref{fig:msd}. The MSD is defined as

@@

\begin{figure}
  \includegraphics[width=0.48\textwidth]{images/msd}
  \caption{\label{fig:msd} (color online) Mean Square Displacement(MSD) vs. time under various temperature measured through ab-initial molecular dynamics. The continuous increasing of MSD at 900K means a phase transition between 800K and 900K.
  }
\end{figure}

@@

However, the structure starts melting when heated to 900 K. Thus we only concentrated on the trusted temperature range ($300K - 800K$), and choose 300 K as a typical temperature to perform the thermal conductivity calculations.

To obtain reliable results of thermal conductivity, the convergence dependence to q points density has to be tested. FIG.\ref{fig:convergence}a shows the convergence of thermal conductivity on $64 \times N \times N$ q points with N increase from 2 to 8. The TC shows its convergence at N=4, above witch the result changes a little but not the computational effort, so we choose the number of transverse q points to be both 4. Besides, the convergence on $N_x \times 4 \times 4$ q points with N increase from 16 to 1024 is also tested as in FIG.\ref{fig:convergence}b. The result of $N_x=200$ could be a good estimate of the converged result and we choose it to calculate TC only for the trends of TC on other properties like temperature et. al to save computer resources. FIG.\ref{fig:convergence}c shows the convergence on phonon cut off free path. The exact convergence could not be obtained even at $N_x=1024$ due to the abnormally large lifetime of long wave phonons show in FIG.\ref{fig:convergence}d, which always exist in low dimensional materials. However, a lot of research shows the relation of TC and phonon cut off free path obeys the rules of \cite{Schelling2002}

@@

and FIG.\ref{fig:convergence}e verifies that it also make sense in bulk \bii and the calculated results are over estimated. As the conclusion, the calculated results of thermal conductivity are not converged even in extremely dense q points, however they can be used as upper limited estimates of the exact results.

\begin{figure}[b]
  \includegraphics[width=0.48\textwidth]{images/convergence}
  \caption{\label{fig:convergence} (color online) Thermal conductivity(TC) convergence vs. q points of bulk \bii. a) Thermal conductivity convergence with $64 \times N \times N$ q points with N increase from 2 to 8. Although not obviously a convergence is reached, the result of N=4 is proved to be a reasonable approximation of convergent result. b) Thermal conductivity convergence with $N_x \times 4 \times 4$ q points with N increase from 16 to 1024. The convergence along the stripe direction is difficult to obtain because of the phonon life time divergence nature of 1D materials and the unreachable mean free path of them.
  }
\end{figure}

The thermal conductivity of \bii is extremely low and anisotropic as shown in FIG.\ref{fig:kappa}. The TC values decreases with increasing temperature and obeys the rule of $\~1/T$, and the largest value at 300K is much less than $1 W/mK$. The TC along the stripes are almost 3 times that of the transverse value which may result from the nature the quasi-1D structure.

@@

where N is the total number of atoms and $\epsilon_{k\sigma} (s)$ is the complex amplitude of atom s for eigen-mode $k\sigma$. The participation ratio presents the fraction of atoms participating in a given mode and effectively indicates the localized modes with $O(1/N)$ and delocalized modes with $O (1)$. It can provide a more detailed information about the localization effect to each mode. The eigenvectors and frequencies are obtained using Phonopy\cite{Togo2008} with $3\times2\times2$ supercell and $15\times15\times15$ mesh sampling. FIG.\ref{fig:phonon} shows the participation ratio of \bii. A lot of phonons have participation ratio below 0.5 which means they are strongly localized and lost the capability to carry energy and they have little contribution to thermal conductivity.

\begin{figure}[b]
  \includegraphics[width=0.48\textwidth]{images/kappa_t}
  \caption{\label{fig:kappa} (color online)Thermal conductivity vs. temperature along the three lattice vectors. The TCs are found to be pretty low. The highly anisotropic structure result in the understandable anisotropic but unexpectedly, the TCs along the stripe and along the inter-stripe direction y are pretty close while the TC along the other inter-stripe direction z is much higher.
  }
\end{figure}

Large anharmonic would be another key factor that account for the low thermal conductivity. FIG.\ref{fig:gv} shows the group velocity of the phonons. Although the frequencies are pretty low (phonons are pretty soft), the group velocities are not so small witch can be understood by the steep phonon dispersion, for example, along the path $\Gamma M$. The maximum group velocity is comparable to that of graphene witch possesses the largest thermal conductivity so far. So group velocity is not responsible for the low TC. The large Gruneisen parameters\cite{Hasegawa1980} showed in FIG.\ref{fig:phonon}b means pretty large anharmonic effects in \bii which suggest strongly three-phonon scattering effects\cite{Hungary1985Phonon}. FIG.\ref{fig:phonon}a shows the three-phonon scattering relaxation time of all the phonons. The mean free path are also show in FIG.7b.Except for the long wave ones, all the life time are pretty small and this is exactly the origin of the low TC. The scattering strength show in FIG.\ref{fig:phonon}a also shows large scattering of phonons.

\begin{figure}[b]
  \includegraphics[width=0.48\textwidth]{images/gv}
  \caption{\label{fig:gv} (color online) Phonon group velocity of bulk \bii. Although the frequencies are pretty low (phonons are pretty soft), the group velocities are not so small witch can be understood by the steep phonon dispersion, for example, along the path $\Gamma M$. The maximum group velocity is comparable to that of graphene witch possesses the largest thermal conductivity so far.
  }
\end{figure}

Low TC means the potential of applications in thermoelectric field and \bii is predicted to be a good thermoelectric material. FIG.\ref{fig:phonon}a shows Seebeck coefficient S at different temperatures. The typical values are in the range of several mV/K and are comparable with that of SnSe. Seebeck coefficient are sensitive to temperature while FIG.\ref{fig:phonon}b shows that electric conduction $\sigma$ are not. The power factor $S^2 \sigma$ are shown in FIG.\ref{fig:phonon}c. Although Seebeck coefficient is smaller at 700K, power factor is higher because of the rise and fall of $\sigma$ around fermi level. However $\kappa{_el}$ is also higher at 700K which eliminate the difference. The electric thermal conductivity is comparable at all the temperature and is even higher, which dominant at high temperature as shown in FIG.\ref{fig:phonon}a. At this point all the factors of figure ZT of merit are collected and the final results are shown in FIG.\ref{fig:phonon}b. After some n-doping or p-doping, ZT reaches its maximum and the best values could be obtained at high temperature. At 700K this ZT value is around 0.8 and this is not bad for the application of thermoelectric energy translation.
